Bulk Semantic Annotation with an Partially Known Knowledge Base

### 2024-11-13 doc/cicm25-sysdesc/sysdesc.tex
more

As \srf addresses mostly power-users, the UI of \srf is command-line based, much like the
early \textsf{ispell} system, see

@@

The step-through

>>>>

As \srf addresses mostly power-users, the UI of \srf is command-line based, much like the
early \textsf{ispell} system, see\ednote{continue}

@@

In our experience, the step-through workflow for annotating term references for symbols
from the domain model is almost an order of magnitude more efficient than writing the
annotations and imports by hand, even for annotators who are familiar with the domain
model. For annotators unfamiliar with the domain model, the unassisted annotation task is
almost infeasible, and the average unfamiliarity naturally grows with the domain
model. The symbol disambiguation process -- on average a word induces $n$\ednote{MK@FS: we
  should try to measure this: for every verbalization (i.e. symbol/word pair) we should
  compute the ratio of those words with the same stem.} choices is still manageable,
requires considerable concentration and domain knowledge, but little knowledge of the
domain model flexiformalization.

The command-line interface is simple and responsive and gives all the necessary
information in a single glance if underlying shell area exceeds ca. $80\times 35$
characters. It is very much geared towards annotating existing documents with respect to a
relatively complete -- pre-existing -- domain model. Otherwise we have to skip too many
terms that should ultimately be annotated and annotation efficiency suffers. This is
currently the case for all non-English languages in the \sTeX commons. Coverage in German
is only about half of that of English and we can already see the practical effects. We
have also experimented with a Slovene introductory math book, and it seems clear that
apart from having to introduce a Slovene stemmer\ednote{MK@FS: there is one for python at
  \url{https://repo.ijs.si/pboskoski/slo_stemmer}}, we would need to manually annotate all
definienda in the book before we can harvest the symbol/verbalization pairs which are a
prerequisite for annotation.

The current interface is not well-suited for on-the-fly annotation while editing while
authoring. For that, the underlying information (the symbol/verbalization pairs harvested
from the domain model) can be integrated into any IDE. In fact we plan to do this for the
next version of the sTeX plugin for VSCode \cite{sTeX-IDE:git}.
